A Plasma Torus Around a Young Low-Mass Star

### 2025-04-03 paper/ms.tex
update

Given that a full radiative transfer simulation was beyond our scope,
we opted to construct a phenomenological model aimed at capturing 
the emission from the circumstellar material.  We did this by
fitting each spectral epoch with a multi-component
gaussian, after having subtracted the time-average line profile as in
Figure~\ref{fig:spec}e.  The results of this exercise are shown in
Figure~\ref{fig:halphamodel}; details in its implementation and
interpretation follow.

To the implement the model, we assumed that the ``inner'' ($K_{\rm
inner}$$\approx$2.5\,$v_{\rm eq}$) clump would be well-fit by a sum of
two gaussians because it visually appears double-peaked in the raw
data from $\phi$=0.15-0.35 and $\phi$=0.65-0.85
(Figure~\ref{fig:spec}b).  We assumed the ``outer'' ($K_{\rm
outer}$$\approx$3.9\,$v_{\rm eq}$) clump would be better fit by a
single gaussian, based on its behavior from $\phi$=0.6-0.9.  Each
gaussian component has three free parameters at each spectral epoch:
the mean $\mu$, standard deviation $\sigma$, and amplitude $A$.
Denote the two gaussians for the inner component $i=\{ 0, 1 \}$,
and the single outer component with $i=2$.  Given the complexity of the
line profile data (Figure~\ref{fig:halphamodel} left column), the
likelihood function for this model is multimodal.  We therefore
imposed the prior constraints that $A_i \sim \mathcal{U}[0,1]$,
$\sigma_i/v_{\rm eq} \sim \mathcal{U}[0,1]$, and further assumed
$K_i(t) \sim
\mathcal{U}[
  K_{\rm inner}\sin(\phi(t)) - v_{\rm eq},
  K_{\rm inner}\sin(\phi(t)) + v_{\rm eq}
]$
for the inner two components, and
$K_i(t) \sim
\mathcal{U}[
  K_{\rm outer}\sin(\phi(t) + \pi) - 2v_{\rm eq},
  K_{\rm outer}\sin(\phi(t) + \pi) + 2v_{\rm eq}
]$
for the outer component.
This latter assumption constrains the mean velocity of each component
to be near the value suggested by visual inspection, but provides
sufficient flexibility for the model to deviate from the assumption of
a sinusoid if the data dictate such behavior.
We fitted each component to the data {\it independently} using scipy's
non-linear least squares \texttt{curve\_fit} method
\cite{Virtanen2020}, and scaled the resulting parameter covariance
matrix by a constant factor to match the sample variance of the
residuals.
The choice to fit the $i=\{ 0, 1 \}$, and $i=2$ components
independently generally did not affect the results compared to if we
have fit all three simultaneously, except for at $\phi=1.00$ (see
Figure~\ref{fig:halphamodel}).

@@

   & Mean1 & Mean2 & Mean3 & Sigma1 & Sigma2 & Sigma3 & Amplitude1 & Amplitude2 & Amplitude3 \\
  BTJD &  &  &  &  &  &  &  &  &  \\

@@

  \caption{Multi-gaussian model fit results to H$\alpha$ line profiles.}

>>>>

Given that a full radiative transfer simulation was outside our scope,
we opted to construct a phenomenological model aimed at capturing 
the emission from the circumstellar material.  We did this by
fitting each spectral epoch with a multi-component
gaussian, after having subtracted the time-average line profile as in
Figure~\ref{fig:spec}e.  The results of this exercise are shown in
Figure~\ref{fig:halphamodel}; details in its implementation and
interpretation follow.

To the implement the model, we assumed that the ``inner'' ($K_{\rm
inner}$$\approx$2.5\,$v_{\rm eq}$) clump would be well-fit by a sum of
two gaussians because it visually appears double-peaked in the raw
data from $\phi$=0.15-0.35 and $\phi$=0.65-0.85
(Figure~\ref{fig:spec}b).  We assumed the ``outer'' ($K_{\rm
outer}$$\approx$3.9\,$v_{\rm eq}$) clump would be better fit by a
single gaussian, based on its behavior from $\phi$=0.6-0.9.  Each
gaussian component has three free parameters at each spectral epoch:
the mean $\mu$, standard deviation $\sigma$, and amplitude $A$.
Denote the two gaussians for the inner component $i=\{ 0, 1 \}$,
and the single outer component with $i=2$.  Given the complexity of the
line profile data (Figure~\ref{fig:halphamodel} left column), the
likelihood function for this model is multimodal.  We therefore
imposed the prior constraints that $A_i \sim \mathcal{U}[0,1]$,
$\sigma_i/v_{\rm eq} \sim \mathcal{U}[0,1]$, and further assumed
$K_i(t) \sim
\mathcal{U}[
  K_{\rm inner}\sin(\phi(t)) - v_{\rm eq},
  K_{\rm inner}\sin(\phi(t)) + v_{\rm eq}
]$
for the inner two components, and
$K_i(t) \sim
\mathcal{U}[
  K_{\rm outer}\sin(\phi(t) + \pi) - 2v_{\rm eq},
  K_{\rm outer}\sin(\phi(t) + \pi) + 2v_{\rm eq}
]$
for the outer component.  This latter assumption mitigates the
multimodality of the likelihood by requiring the mean velocity of each
component to be within a one or two $v_{\rm eq}$ of the time-variable
sinusoid suggested by visual inspection.  We fitted each component to
the data {\it independently} using scipy's non-linear least squares
\texttt{curve\_fit} implementation \cite{Virtanen2020}, and scaled the
resulting parameter covariance matrix by a constant factor to match
the sample variance of the residuals.  The resulting means,
amplitudes, and standard devations for each component are given in
Extended Data Table~\ref{tab:halphamodelparams}.

Caution is required in the interpretation of this model's results.
The model at times overfits, because it targets the most significant
spectral feature around the priors discussed above, and sometimes
these features are not statistically significant.  During such
spectral epochs, e.g.~the ``outer'' clump at $\phi$=1.07, the model is
fitting noise, not signal.  
At times the model underfits.  For instance, the sudden blue rise near
$\phi$=0.2 is not well-matched by the functional form of a gaussian;
the assumed functional form is one of convenience.
A final point of caution is that the model fits the $i=\{ 0, 1 \}$,
and $i=2$ components independently.  At $\phi=0.0$ and $1.0$,
Figure~\ref{fig:halphamodel} suggests that this means the emission might
hypothetically be from either the inner or outer components.  A
physical interpretation however is that this is the phase during which
the inner component is in transit, and the outer is in secondary
occultation and blocked by the star.  At $\phi=0.0$, the double-peaked
emission profile also matches that seen shortly afterward when the
inner clump is viewed off-disk.  A physical interpretation of the
model should therefore discard the results at this particular epoch
for the outer clump.

Taking these caveats into account, the results from Extended Data
Table~\ref{tab:halphamodelparams} can be used to evaluate various
models for the orbits.

, and in particular to assess their periods and
semi-amplitudes.

@@

  BTJD & $\mu_0$ & $\mu_1$ & $\mu_2$ & $\sigma_0$ & $\sigma_1$ & $\sigma_2$ & $A_0$ & $A_1$ & $A_2$ \\

@@

  \caption{Multi-gaussian model fit results to H$\alpha$ line
  profiles.  The results should be treated with caution; only a
  subset of spectral epochs provide statistically significant
  detections of the circumstellar emission (see Extended Data
  Figure~\ref{fig:halphamodel}).}
